AA: the Algorithmic Auto-regulation (Software Development) Methodology

### 2013-06-01 paper.tex
adicionado Penalva nos autores e 2 bibs
some minor grammatical corrections

One of the defining features of modern times is the widening geographical
distribution of software teams~\cite{last2003}. This is responsible for the so
called Global Software Development (GSD)~\cite{german2003}.  An example is the
free software movement. Projects and institutions like Mozilla Foundation has
several employees and thousands of voluntary developers distributed across many
countries. The same is true for GNOME~\cite{german2003}, OpenBSD, MySQL or
Apache Software Foundation, to cite just a few\footnote{Ohloh, the open source
  network, have a more complete and constantly updated list of the most active
  projects on-line at \url{www.ohloh.net}}. Along the free and open software
projects, GSD has a growing popularity in every niche of the software industry
as a whole, even on those distributing their software with proprietary
licenses. This phenomenon is attributed to a variety of factors such as a larger
labor pool, natural globalization of software companies and foundations or even
the premise of cheaper cost of production~\cite{komi2005}.

@@

There has been a large amount of research in methodologies to
deal with distributed teams of developers. Although this article focuses in GSD,
some principles involved on those methodologies could be used on smaller teams
of developers working in the same place, time period and with minor cultural
differences. Moreover, ``distributed development'' is generally tough as being
global which is not true. AA was applyed to a team that work at the
same city but on different times (more details on Section~\ref{results}). Even
smaller groups of developers working on the same building could use GSD
methodologies. A thorough survey of these methodologies is beyond the scope of
this paper. This section presents a brief overview.

Various methodologies for GSD were built around the factors that affect
distributed team works. As proposed by Carmel~\cite{carmel1999}, these comprises
three distances: geographical, cultural and temporal. Geographical distance
prejudice \emph{coordination}, being the act of integrating all the tasks
distributed between units~\cite{carmel2001}; \emph{control}, or the process to
maintain specific goals, policies or quality levels; and
\emph{communication}. All these factors are correlated, for example, a team needs
to have good communication to work on tasks related to a specific problem. Cultural
distance encompass differences on organizational and natural culture. Spoken
language, unit and ethnic values are common forms of this distance. Some
companies prefer to situate development units in foreign locations with minimal
cultural distance (e.g.\ an American firm prefer Ireland, because of spoken
language similarity~\cite{carmel2001}). Finally, the temporal distance
prejudices synchronous communications like telephone or video-conferences.
Developers working on different time-zones manage
their agenda guided by this temporal distance.

@@

While considering free software projects instead of companies, similar factors
are present and some methodologies arises. German~\cite{german2003} gives a
concise review of methodologies used by the GNOME project, one of the most
active free software projects and used by companies like Sun Microsystems. It is
interesting to note a difference on the viewpoint present on his paper: German
focuses the methodology description on code. He start explaining that GNOME is
separated into modules (76 on version 2.4, to be precise) and each module has
one maintainer who divide his modules into separated parts in which other
developers can work on independent tasks, along other responsibilities. As like
others free software projects, all the development was done using a
\emph{bugtracker} to bugs and issues management, mail lists and Internet Relay
Chat (IRC) to discussion and communication and a software version
management like SVN or Git. He finishes by observing that periodical, e.g. yearly, conferences like GUADEC
is common on free and open source projects for face-to-face meetings and is based
each time on a different place.

@@

This paper has presented a simple methodology to GSD, being the development conducted
on large or small groups of software developers, working on different countries
or even at the same room. The AA methodology implements a simple system where
each developer takes notes of his work, generating a periodical log of small text
sentences. The sum of those sentences, along an entire session of work, results
in a complete report. The report is made public available through a Website and
are validated by other developers sorted aleatory by the AA Web server.

>>>>

One of the defining features of modern times is the widening geographical
distribution of software teams~\cite{last2003}. This is responsible for the so
called Global Software Development (GSD)~\cite{german2003,Fryer,Begel}.  An example is the
free software movement. Projects and institutions like Mozilla Foundation has
several employees and thousands of voluntary developers distributed across many
countries. The same is true for GNOME~\cite{german2003}, OpenBSD, MySQL or
Apache Software Foundation, to cite just a few\footnote{Ohloh, the open source
  network, have a more complete and constantly updated list of the most active
  projects on-line at \url{www.ohloh.net}}. Along the free and open software
projects, GSD has a growing popularity in every niche of the software industry
as a whole, even on those distributing their software with proprietary
licenses. This phenomenon is attributed to a variety of factors such as a larger
labor pool, natural globalization of software companies and foundations or even
the premise of cheaper cost of production~\cite{komi2005}.

@@

There has been a large amount of research in methodologies to
deal with distributed teams of developers. Although this article focuses in GSD,
some principles involved on those methodologies could be used on smaller teams
of developers working in the same place, time period and with minor cultural
differences. Moreover, ``distributed development'' is generally tought as being
global which is not true. AA was applyed to a team that work at the
same city but on different timings (more details on Section~\ref{results}). Even
smaller groups of developers working on the same building could use GSD
methodologies. A thorough survey of these methodologies is beyond the scope of
this paper. This section presents a brief overview.

Various methodologies for GSD were built around the factors that affect
distributed team works. As proposed by Carmel~\cite{carmel1999}, these comprises
three distances: geographical, cultural and temporal. Geographical distance
prejudice \emph{coordination}, being the act of integrating all the tasks
distributed between units~\cite{carmel2001}; \emph{control}, or the process to
maintain specific goals, policies or quality levels; and
\emph{communication}. All these factors are correlated, for example, a team needs
to have good communication to work on tasks related to a specific problem. Cultural
distance encompass differences in organizational and natural culture. Spoken
language, unit and ethnic values are common forms of this distance. Some
companies prefer to situate development units in foreign locations with minimal
cultural distance (e.g. an American firm prefer Ireland, because of spoken
language similarity~\cite{carmel2001}). Finally, the temporal distance
prejudices synchronous communications like telephone or video-conferences.
Developers working on different time-zones manage
their agenda guided by this temporal distance.

@@

While considering free software projects instead of companies, similar factors
are present and some methodologies arises. German~\cite{german2003} gives a
concise review of methodologies used by the GNOME project, one of the most
active free software projects and used by companies like Sun Microsystems. It is
interesting to note a difference in the viewpoint present on his paper: German
focuses the methodology description on code. He start explaining that GNOME is
separated into modules (76 on version 2.4, to be precise) and each module has
one maintainer who divide his modules into separated parts in which other
developers can work on independent tasks, along with other responsibilities. As like
others free software projects, all the development was done using a
\emph{bugtracker} to bugs and issues management, mail lists and Internet Relay
Chat (IRC) to discussion and communication and a software version
management like SVN or Git. He finishes by observing that periodical, e.g. yearly, conferences like GUADEC
are common in free and open source projects for face-to-face meetings and are based
each time on a different place.

@@

This paper has presented a simple methodology to GSD, being the development conducted
on large or small groups of software developers, working in different countries
or even at the same room. The AA methodology implements a simple system where
each developer takes notes of his work, generating a periodical log of small text
sentences. The sum of those sentences, along an entire session of work, results
in a complete report. The report is made public available through a Website and
are validated by other developers sorted aleatory by the AA Web server.
